### 2020-06-26 manuscript/manuscript.rmd
some chunks that are not validated with code to verify assertions in text

As others have noted (XX), our summary of previous phylogenetic analyses (Fig. 3A) show that Ctenophora-sister is supported over analyses that span the full range of outgroup sampling and models used to date, including some analyses with restricted outgroup sampling and models that accommondate site-heterogeneous equilibrium frequencies with CAT. The only analyses that support Porifera-sister ahveh reduced outgroup sampling (Choanimalia, Holozoa) and site heterogeneous models with CAT. Model adequacy assessments generally favor GTR+CAT over Poisson+CAT or site-homogeneous models [@pisani2015genomic; @feuda2017improved], but because CAT is so parameter rich, many analyses that use a model with CAT do not converge. Some of the unconvered analyses have an unresolved animal root. The fact that Porifera-sister is recovered only for particular models with particular outgroup sampling indicates that model and outgroup interact, and that this interaction is fundamental to understanding the range of results obtained across analyses.

@@

One of the challenges of interpreting support for the placement of the animal root is that different programs, software versions, and settings have been used across studies, and phylogenetic analysis decisions have been approached in very different ways. Thus, it easily causes confusions with the increase number of studies.  Here we first reanalyze the primary matrices from each study under consistent conditions with IQ-TREE [@nguyen2014iq] under a panel of evolutionary models. We selected this tool because it has greater model flexibility, computational time and accuracy than other tools in maximum likelihood (ML) framework [@zhou2017evaluating] (Fig. 3B; Supplementary Table 3). Importantly, we also include a CAT-like, site-heterogeneous C10 -C60 (C60) models [@si2008empirical] that are implemented in both ML framework in our analyses and it is interesting to compare results between C60 model to CAT models since it has never been used in any matrices analyzed here.

Here we reanalyzed 17 matrices from 13 studies that are used as main conclusions from their original paper (Extended Data Table 1). We also progressively trimmed each data matrix with three level of outgroup choice if applicable, original outgroup sampling permitting (Fig 1C). We first tested a variety of models for each matrix and compared the relative fit of site-homogeneous and site-heterogeneous C60 models using ModelFinder [@kalyaanamoorthy2017modelfinder] in IQ-TREE (except only LG+C60 models are used in 3 large matrices). We found in all cases, WAG/LG+C60 models fit these matrices better than the site-homogeneous models under BIC criteria and we then inferred support under the best-fit model from IQ-TREE (Extended Data Table 1). We then analyzed every matrix under a panel of standard site-heterogeneous and site-homogeneous models, including WAG, GTR and Poisson+C60. Moreover, we also used the best-fit model identified above with the removal of all distantly related outgroups and only include Choanoflagellata as outgroups. Interestingly, with the exception of Moroz2014_3d and all Nosenko2013 matrices that supports neither Porifera-sister or Ctenophora-sister, all analysis conducted herein in IQ-TREE strongly supported the ctenophore-sister hypothesis, even with CAT-like, site-heterogeneous C60 models (Poisson+C60 and WAG/LG+C60) with the most closely related outgroups (Fig. 3B; Supplementary Table 3).

>>>>

```{r eval=FALSE}

# Code to double check and explore some of the assertions in the text

# Simion2017 is the only Holozoa analysis to support Porifera-sister
analyses_processed %>% filter(status=="published") %>% filter(clade=="Holozoa") %>% filter(result=="Porifera-sister")

# Check out the CAT analyses that supported Ctenophora sister
analyses_processed %>% filter(status=="published") %>% filter(result=="Ctenophora-sister") %>% filter( grepl("CAT", model))

Our summary of previous phylogenetic analyses (Fig. 3A) show that Ctenophora-sister is supported over analyses that span the full range of outgroup sampling and models used to date, including some analyses with restricted outgroup sampling and models that accommondate site-heterogeneous equilibrium frequencies with CAT. This is consistent with previous assessments of the problem (XX), but draws on a more systematic overview. The only analyses that support Porifera-sister ahveh reduced outgroup sampling (Choanimalia, Holozoa) and site heterogeneous models with CAT. Model adequacy assessments generally favor GTR+CAT over Poisson+CAT or site-homogeneous models [@pisani2015genomic; @feuda2017improved], but because CAT is so parameter rich, many analyses that use a model with CAT do not converge. Some of the unconvered analyses have an unresolved animal root. The fact that Porifera-sister is recovered only for particular models with particular outgroup sampling indicates that model and outgroup interact, and that this interaction is fundamental to understanding the range of results obtained across analyses.

@@

One of the challenges of interpreting support for the placement of the animal root is that different programs, software versions, and settings have been used across analyses. This extensive variation can make it difficult to identify the primary factors that lead to different results. Here we first reanalyze the primary matrices from each study under consistent conditions with IQ-TREE [@nguyen2014iq] under a panel of evolutionary models. We selected this tool because it has greater model flexibility than other maximum likelihood (ML) tools [@zhou2017evaluating] (Fig. 3B; Supplementary Table 3). Importantly, it has C10-C60 equilibrium frequencies [@si2008empirical] that, like CAT, allow models to accommodate heterogeneity across sites. There are some implementation differences. For example, CAT estimates the number of distinct site categories, the assigment of sites to those categories, and the equilibrium frequencies of each of those categories, while the C models specify the number of categories *a priori* (the C10 model as 10 categories, etc...).

```{r eval=FALSE}

# Code to double check some of the assertions in the text

# See results (and other fields) for new Cxx analyses, all are Ctenophora-sister with exception of Moroz2014_3d and Nosenko matrices
analyses_processed %>% filter(status=="new") %>% filter( grepl("C\\d\\d",model) ) %>% print(n=40)

For each of the published studies, we selected the matrix that was the primary focus of the manuscript for further analysis. For each of these matrices, we progressively trimmed taxon sampling of to create Opisthokonta, Holozoa, and Choanimalia versions, where permitted by original outgroup sampling. This produced `r analyses_processed %>% filter(status=="new") %>% nrow()` matrices. Each of these was then analyzed using several models. We used modelfinder to identify the

We tested a variety of models for each matrix and compared the relative fit of site-homogeneous and site-heterogeneous C60 models using ModelFinder [@kalyaanamoorthy2017modelfinder] in IQ-TREE (except only LG+C60 models are used in 3 large matrices). We found in all cases, WAG/LG+C60 models fit these matrices better than the site-homogeneous models under BIC criteria and we then inferred support under the best-fit model from IQ-TREE (Extended Data Table 1). We then analyzed every matrix under a panel of standard site-heterogeneous and site-homogeneous models, including WAG, GTR and Poisson+C60. Moreover, we also used the best-fit model identified for Choanoflagellata taxon sampling. Interestingly, with the exception of Moroz2014_3d and all Nosenko2013 matrices that supports neither Porifera-sister or Ctenophora-sister, all analysis conducted herein in IQ-TREE strongly supported the ctenophore-sister hypothesis, even with CAT-like, site-heterogeneous C60 models (Poisson+C60 and WAG/LG+C60) with the most closely related outgroups (Fig. 3B; Supplementary Table 3).
